### 2025-11-24 article/fig1.tex

\documentclass[border=4pt]{standalone}
\usepackage{graphicx}

\begin{document}

% Colonne gauche
\begin{minipage}{0.48\linewidth}
  \centering
  \includegraphics[width=\linewidth]{../outputs/hist/dispo/quotidien/hydro/hist_len_years_cum.pdf}

  {\scriptsize Stations at daily time step (n = 8198)}
\end{minipage}
\hfill
% Colonne droite
\begin{minipage}{0.48\linewidth}
  \centering
  \includegraphics[width=\linewidth]{../outputs/hist/dispo/horaire/hydro/hist_len_years_cum.pdf}

  {\scriptsize Stations at hourly time step (n = 2315)}
\end{minipage}

\end{document}